Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2023-01-10 edubots_surveys.Rnw
Update edubots_surveys.Rnw
Co-authored-by: Juan Julián Merelo Guervós <jjmerelo@gmail.com>

In this study, data collection involved two phases:
 \begin{itemize}
 \item The first phase aimed to collect feedback from students about their
   students’ preferences of tools/applications for chatting and messaging, how
   they use them in educational context, who they like to be with in the class
   messaging groups, and their expectations of chatbots in assisting them during
   their learning process. Therefore, two surveys for bachelor and master degree
   students at the University of Granada (Spain) were designed and answers from
   more than 250 students were collected. The key results from the student
   surveys indicate a preference for using familiar messaging applications, such
   as Telegram or WhatsApp. Students also expressed interest in using chatbots
   that can assist them with organizing their course schedules, help them access
   their assignment grades and facilitate searching for resources.
\item The results of the first phase were fed into the second phase of the
  study, which focused on teachers’ opinions about using chatbots and messaging
  applications in classes. These results are the main focus of this paper.
\end{itemize}

Initially, a single survey was developed for this purpose and responses from 300
higher education teachers were collected and analysed; this
  will be detailed in Section \ref{sec:meth}. The responses' analysis led to
the preparation of a second survey for teachers and 200 responses were
collected, again following the methodology that will be
  explained later. The questions of the second survey focused on asking about
  teachers' needs, opinions and preferences of the development of future
  technology-enhanced tools and its potential impact on educational institutions
  policies. In order to assess these, this paper layed out the following
  research questions:

@@

The widespread and rapid adoption of free Mobile Instant Messaging
(MIM) tools/platforms such as WhatsApp, Telegram, WeChat and Facebook
Messenger stems from their simplicity, ease of use and multi-modality
(i.e. video, audio, text) \cite{tang2017mobile}. Using these tools in
higher education institutions facilitates the delivery of personalized learning
that occurs anytime anywhere, and promotes collaborative learning
experiences and group discussions \cite{panah2020study}.

@@

In our survey, we could not find an answer about the right moment to introducing chatbot technologies. The response to questions regarding to change of
behavior during and after the first stay-at-home stage of the pandemic do not suggest that a major (or minor, such as new higher education laws\footnote{We seem to have one of these, at least in Spain, every 10 years or so.} or the introduction of new degrees) crisis could be an opportunity to introduce new technologies. The reason being that it does not
bring major changes in attitude 
%(as a minor ``crisis'' we could mention any institutional event that brings about methodological changes, such as new higher education laws\footnote{We seem to have one of these, at least in Spain, every 10 years or so.}), 
%or simply the introduction of new degrees. 
Although changes driven by external forces do offer the chance of piloting new technologies, they do not
seem to bring internal changes in educators' attitudes (which, after all, is
bound to be the same). In absence of a clear answer in this direction, the right
moment to introduce new technologies is always when the IT and managing staff is
ready to support it (since, as we have seen before, ``official'' support is one of
the factors that is most valued by teachers).

>>>>

In this study, data collection involved two phases:
 \begin{itemize}
\item The first phase aimed to collect feedback from students about their
  students’ preferences of tools/applications for chatting and
  messaging, how they use them in educational context, who they like to be
  with in the class messaging groups, and their expectations of chatbots
  in assisting them during their learning process. Therefore, two surveys
  for bachelor and master degree students at the University of Granada (Spain)
  were designed and answers from more than 250 students were collected. The key
  findings of the students' surveys reveal that they prefer to use messaging
  familiar applications, such as Telegram or WhatsApp.
  Students also expressed interest in chatbots that can assist them with
  organizing their course schedules, help them access their assignment grades
  and facilitate searching for resources.
\item The results of the first phase were fed into the second phase of the
  study, which focused on teachers’ opinions about using chatbots and messaging
  applications in classes. These results are the main focus of this paper.
\end{itemize}

Initially, a single survey was developed for this purpose and responses from 300
higher education teachers were collected and analysed\textcolor{black}{; this will be detailed in Section \ref{sec:meth}}. The responses' analysis
led to the preparation of a second survey for teachers and 200 responses were
collected\textcolor{black}{, again following the methodology that will be explained later. The questions of the second survey focused on asking about teachers'
needs, opinions and preferences of the development of future technology-enhanced
tools and its potential impact on educational institutions policies. In order to assess these, this paper layed out the following research questions}:

@@

% The Use of Mobile Instant Messaging (MIM) Platforms in Higher Education
The widespread and rapid adoption of free Mobile Instant Messaging
(MIM) tools/platforms such as WhatsApp, Telegram, WeChat and Facebook
Messenger stems from their simplicity, ease of use and multi-modality
(i.e. video, audio, text) \cite{tang2017mobile}. Using these tools in
higher education institutions facilitates the delivery of personalized learning
that occurs anytime anywhere, and promotes collaborative learning
experiences and group discussions \cite{panah2020study}.

@@

% Only some COVID-response relationship.
In our survey, we could not find an answer about the right moment to introducing chatbot technologies. The response to questions regarding to change of
behavior during and after the first stay-at-home stage of the pandemic do not suggest that a major (or minor, such as new higher education laws\footnote{We seem to have one of these, at least in Spain, every 10 years or so.} or the introduction of new degrees) crisis could be an opportunity to introduce new technologies. The reason being that it does not
bring major changes in attitude 
%(as a minor ``crisis'' we could mention any institutional event that brings about methodological changes, such as new higher education laws\footnote{We seem to have one of these, at least in Spain, every 10 years or so.}), 
%or simply the introduction of new degrees. 
Although changes driven by external forces do offer the chance of piloting new technologies, they do not
seem to bring internal changes in educators' attitudes (which, after all, is
bound to be the same). In absence of a clear answer in this direction, the right
moment to introduce new technologies is always when the IT and managing staff is
ready to support it (since, as we have seen before, ``official'' support is one of
the factors that is most valued by teachers).
